---
title: "EDA"
author: "Donghyun Lee"
date: "2024-05-26"
output:
  pdf_document:
    latex_engine: xelatex
mainfont: NanumGothic
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
```

### 1. 잠실역

```{r, warning=FALSE, message=FALSE}
df <- read_csv("data/df_jamsil.csv")
```


### temp_avg

```{r}
ggplot(data=df, mapping=aes(x=temp_avg, y=total)) +
  geom_point() +
  labs(title = "Average Temperature")
cor(df$temp_avg, df$total)
```

\newpage
### precipitation

```{r}
ggplot(data=df, mapping=aes(x=precipitation, y=total)) +
  geom_point() +
  labs(title = "Precipitation")
cor(df$precipitation, df$total)
```

\newpage
### peak_jamsil

```{r}
df$peak_jamsil <- factor(df$peak_jamsil)
ggplot(data=df, mapping=aes(x=peak_jamsil, y=total)) +
  geom_boxplot() +
  labs(title = "Cherry blossom, Children's day, Christmas")
```

\newpage
### holiday

```{r}
df$holiday <- factor(df$holiday)
ggplot(data=df, mapping=aes(x=holiday, y=total)) +
  geom_boxplot() +
  labs(title = "Lunar New Year, Chuseok")
```

\newpage
### 2. 강남역

```{r, warning=FALSE, message=FALSE}
df <- read_csv("data/df_gangnam.csv")
```


### temp_avg

```{r}
ggplot(data=df, mapping=aes(x=temp_avg, y=total)) +
  geom_point() +
  labs(title = "Average Temperature")
cor(df$temp_avg, df$total)
```
\newpage
### precipitation

```{r}
ggplot(data=df, mapping=aes(x=precipitation, y=total)) +
  geom_point() +
  labs(title = "Precipitation")
cor(df$precipitation, df$total)
```
\newpage
### peak_gangnam

```{r}
df$peak_gangnam <- factor(df$peak_gangnam)
ggplot(data=df, mapping=aes(x=peak_gangnam, y=total)) +
  geom_boxplot() +
  labs(title = "Children's day")
```

\newpage
### holiday

```{r}
df$holiday <- factor(df$holiday)
ggplot(data=df, mapping=aes(x=holiday, y=total)) +
  geom_boxplot() +
  labs(title = "Lunar New Year, Chuseok")
```

\newpage
### 3. 여의도역

```{r, warning=FALSE, message=FALSE}
df <- read_csv("data/df_yeouido.csv")
```

### temp_avg

```{r}
ggplot(data=df, mapping=aes(x=temp_avg, y=total)) +
  geom_point() +
  labs(title = "Average Temperature")
cor(df$temp_avg, df$total)
```
\newpage
### precipitation

```{r}
ggplot(data=df, mapping=aes(x=precipitation, y=total)) +
  geom_point() +
  labs(title = "Precipitation")
cor(df$precipitation, df$total)
```
\newpage
### peak_yeouido

```{r}
df$peak_yeouido <- factor(df$peak_yeouido)
ggplot(data=df, mapping=aes(x=peak_yeouido, y=total)) +
  geom_boxplot() +
  labs(title = "Firework")
```

\newpage
### holiday

```{r}
df$holiday <- factor(df$holiday)
ggplot(data=df, mapping=aes(x=holiday, y=total)) +
  geom_boxplot() +
  labs(title = "Lunar New Year, Chuseok")
```

\newpage
### 4. 서울역

```{r, warning=FALSE, message=FALSE}
df <- read_csv("data/df_seoul.csv")
df <- df %>% mutate(total = 승차총승객수 + 하차총승객수)
```


### temp_avg

```{r}
ggplot(data=df, mapping=aes(x=temp_avg, y=total)) +
  geom_point() +
  labs(title = "Average Temperature")
cor(df$temp_avg, df$total)
```
\newpage
### precipitation

```{r}
ggplot(data=df, mapping=aes(x=precipitation, y=total)) +
  geom_point() +
  labs(title = "Precipitation")
cor(df$precipitation, df$total)
```
\newpage
### peak_seoul_1

```{r}
df$peak_seoul_1 <- factor(df$peak_seoul_1)
ggplot(data=df, mapping=aes(x=peak_seoul_1, y=total)) +
  geom_boxplot() +
  labs(title = "Children's day")
```

\newpage
### peak_seoul_2

```{r}
df$peak_seoul_2 <- factor(df$peak_seoul_2)
ggplot(data=df, mapping=aes(x=peak_seoul_2, y=total)) +
  geom_boxplot() +
  labs(title = "Lunar new year day, Chuseok")
```

\newpage
### peak_seoul_3

```{r}
df$peak_seoul_3 <- factor(df$peak_seoul_3)
ggplot(data=df, mapping=aes(x=peak_seoul_3, y=total)) +
  geom_boxplot() +
  labs(title = "A day before and after Lunar new year day, Chuseok")
```

\newpage
### 5. 고속터미널역

```{r, warning=FALSE, message=FALSE}
df <- read_csv("data/df_express_terminal.csv")
df <- df %>% mutate(total = 승차총승객수 + 하차총승객수)
```

### temp_avg

```{r}
ggplot(data=df, mapping=aes(x=temp_avg, y=total)) +
  geom_point() +
  labs(title = "Average Temperature")
cor(df$temp_avg, df$total)
```
\newpage
### precipitation

```{r}
ggplot(data=df, mapping=aes(x=precipitation, y=total)) +
  geom_point() +
  labs(title = "Precipitation")
cor(df$precipitation, df$total)
```
\newpage
### peak_express_1

```{r}
df$peak_express_1 <- factor(df$peak_express_1)
ggplot(data=df, mapping=aes(x=peak_express_1, y=total)) +
  geom_boxplot() +
  labs(title = "Children's day")
```
\newpage
### peak_express_2

```{r}
df$peak_express_2 <- factor(df$peak_express_2)
ggplot(data=df, mapping=aes(x=peak_express_2, y=total)) +
  geom_boxplot() +
  labs(title = "Lunar new year day, Chuseok")
```
\newpage
### peak_express_3

```{r}
df$peak_express_3 <- factor(df$peak_express_3)
ggplot(data=df, mapping=aes(x=peak_express_3, y=total)) +
  geom_boxplot() +
  labs(title = "A day before and after Lunar new year day, Chuseok")
```
\newpage
### peak_express_4

```{r}
df$peak_express_4 <- factor(df$peak_express_4)
ggplot(data=df, mapping=aes(x=peak_express_4, y=total)) +
  geom_boxplot() +
  labs(title = "The first and last day of the year")
```